---
title: "Next_word_pred_model"
author: "Joe Rubash"
date: "December 12, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---


# plan of attack ("make it work", "make it right", "make it fast")
Stage one - make a simple model using calculated frequencies from n-grams
1. read in data and make a corpus with all sources
2. to start out sample a small portion of this corpus
3. use qunteda package to create unigram (to lower case, strip punc, remove extra 
white space. etc)
4. create a dictionary to assign integer to unique words in unigram
5. build unigram, bigram and trigram document term matrices
6. split ngrams to give previous word(s) and next word (to predict)
7. calc frequencies
8. to colsum of frequencies to
9. create backoff method to start with highest order n-gram then move to lowest
order n-gram if no solution found
10. create method to evaluate accuracy and efficiency of model. Make this easy to 
run so I can later itterate quickly through different ideas to improve


stage two use smoothing to improve prediction and various methods to spee up
8. figure out how to alter frequencies to predict unkowns
9. su

```{r setup}
# clear environment
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)

# required libraries
library(tidyverse)
library(readr)
library(readtext)
library(purrr)
library(quanteda)
library(tm)
library(data.table)
library(stringi)
library(stringr)
library(knitr)
library(rbenchmark)
library(doParallel)
library(R.utils)

# Functions---------------------------------------------------------------------

# Generic function for parallelizing any task (when possible)
parallelizeTask <- function(task, ...) {
  # Calculate the number of cores
  ncores <- detectCores() - 1
  # Initiate cluster
  cl <- makeCluster(ncores)
  registerDoParallel(cl)
  #print("Starting task")
  r <- task(...)
  #print("Task done")
  stopCluster(cl)
  r
}

# create function to read in convert, sample then write to file text data
sample_data <- function(file){
    readLines(con = file,
              skipNul = TRUE,
              encoding="UTF-8") %>%
    sample(.,
           size = sample.size * length(.),
           replace = FALSE)
    }

# Functions to create ngrams and then calculate frequencies
# Create ngrams
create_ngrams <- function(corpus, ngrams){
    tokens(corpus, what = "word",
                              remove_numbers = TRUE,
                              remove_punct = TRUE,
                              remove_symbols = TRUE,
                              remove_separators = TRUE,
                              remove_twitter = TRUE, 
                              remove_hyphens = TRUE,
                              remove_url = TRUE,
                              ngrams = ngrams,
                              skip = 0L,
                              concatenator = "_",
                              verbose = quanteda_options("verbose"),
                              include_docvars = TRUE) %>% 
    tokens_tolower(.) %>% 
    unlist(., use.names = FALSE)
    }

# Model data
# convert to data.table then split bigram into base and prediction
split_base_pred_freq <- function(ngram){
    dt <- data.table(text = ngram)
    dt <- dt[, text := stri_replace_last_fixed(text,"_", " ")]
    dt <- dt[, c("base", "pred") 
      := data.table(stri_split_fixed(text, " ", simplify = TRUE))]
    dt <- dt[, text := NULL]
    dt <- setkey(dt, base)
    dt <- dt[, ":=" (count = uniqueN(pred),
               base.count = .N),
             key = base]
    dt <- dt[count > 1]
    dt <- dt[, prob := count / base.count]
    dt <- dt[ , .SD[which.max(count)],
              key = base]
    return(dt)
}


# Download data-----------------------------------------------------------------
# download and unzip data folder
if(!dir.exists("../data/Coursera-SwiftKey/")){
    file.url <- "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
    temp <- tempfile()
    download.file(file.url, temp, mode = "wb")
    unzip(temp, exdir = "../data/Coursera-SwiftKey")
    rm(temp)
}

# Sample data------------------------------------------------------------------
# define sample size, read in all english data files, sample, combine, add lines and save to file as csv
sample.size <- 0.25 # delete sample files and rerun code if changing sample size

# read in and sample all text data
sample.blogs <- sample_data(file = "../data/Coursera-SwiftKey/final/en_US/en_US.blogs.txt")
sample.news <- sample_data(file = file("../data/Coursera-SwiftKey/final/en_US/en_US.news.txt", "rb"))
sample.twitter <- sample_data(file =  "../data/Coursera-SwiftKey/final/en_US/en_US.twitter.txt")

# Combine all text samples
sample.text <- c(sample.blogs, sample.news, sample.twitter) 

# split sample.text into train, validation and test
set.seed(123)
training <- sample(sample.text, size = 0.90 * length(sample.text))
temp.test <- sample.text[!(sample.text %in% training)]
valid <- sample(temp.test, size = 0.5 * length(temp.test))
test <- temp.test[!(temp.test %in% valid)]

# remove unneeded objects
rm(sample.blogs, sample.news, sample.twitter, temp.test)

```
```{r Create Model}
# the following code seems to work well to create unigram and bigram tokens from 
# combined text. The tokens can then be converted from lists to a single character
# vector.

# Prepare corpus by sentences---------------------------------------------------
corpus.sentences <- corpus(training) %>%
    corpus_reshape(., to = "sentence") 


# Create ngram frequencies data table-------------------------------------------

# unigrams
unigram <- create_ngrams(corpus = corpus.sentences, ngrams = 1L)
unigram.dt <- data.table(text = unigram)
unigram.dt <- unigram.dt[, count := .N, by = text]
unigram.dt <- unigram.dt[, prob := count / length(unigram.dt$count)]
unigram.dt <- unigram.dt[order(count, decreasing = TRUE)]
unigram.dt <- unique(unigram.dt)

# bigrams
# start by initiating all cores for parallel processing
bigram <- create_ngrams(corpus = corpus.sentences, ngrams = 2L)
bigram.freq <- split_base_pred_freq(ngram = bigram)
rm(bigram)

# trigrams
trigram <- create_ngrams(corpus = corpus.sentences, ngrams = 3L)
trigram.freq <- split_base_pred_freq(ngram = trigram)
rm(trigram)

# quadgrams
quadgram <- create_ngrams(corpus = corpus.sentences, ngrams = 4L)
quadgram.freq <- split_base_pred_freq(ngram = quadgram)
rm(quadgram)

# pentagrams
pentagram <- create_ngrams(corpus = corpus.sentences, ngrams = 5L)
pentagram.freq <- split_base_pred_freq(ngram = pentagram)
rm(pentagram, corpus.sentences)

# combine ngrams
ngram.list <- list(bigram.freq, trigram.freq, quadgram.freq, pentagram.freq)
ngrams.freq <- rbindlist(ngram.list)
rm(ngram.list, bigram.freq, trigram.freq, quadgram.freq, pentagram.freq)
ngrams.freq <- setkey(ngrams.freq, base)

# write resulting ngram to file
fwrite(ngrams.freq, file = paste("../data/temp.data/ngrams.freq", sample.size, ".csv"))

getwd()
# Save an object to file under the "next_word_prediction" app
saveRDS(ngrams.freq.15, file = "./Next_word_prediction/model.data/ngrams.freq.15.rds")
# Restore the object
ng.15.test <- readRDS(file = "./Next_word_prediction/model.data/ngrams.freq.15.rds")

# Save an object to file under the "next_word_prediction" app
saveRDS(ngrams.freq.0.25, file = "./Next_word_prediction/model.data/ngrams.freq.25.rds")
# Restore the object
ng.25.test <- readRDS(file = "./Next_word_prediction/model.data/ngrams.freq.25.rds")

```
```{r Read in model}
# read in previously built model from file
ptm <- proc.time() # start time
getwd()
ngrams.freq.15 <- read.csv(file = "./data/temp.data/ngrams.freq.15.csv")
ngrams.freq.15 <- as.data.table(ngrams.freq.15)
ngrams.freq.15 <- setkey(ngrams.freq.15, base)

proc.time() - ptm # end time and difference

ptm <- proc.time() # start time
getwd()
ngrams.freq.0.25 <- read.csv(file = "./data/temp.data/ngrams.freq 0.25 .csv")
ngrams.freq.0.25 <- as.data.table(ngrams.freq.0.25)
ngrams.freq.0.25 <- setkey(ngrams.freq.0.25, base)

proc.time() - ptm # end time and difference
```
```{r prep validation data}
# Prepare corpus by sentences
corpus.sentences <- corpus(valid) %>%
    corpus_reshape(., to = "sentence") 

# Create pentagrams for validation
pentagram <- create_ngrams(corpus = corpus.sentences, ngrams = 5L)

# split pentagram into base and prediction
split_base_pred <- function(ngram){
    dt <- data.table(text = ngram)
    dt <- dt[, text := stri_replace_last_fixed(text,"_", " ")]
    dt <- dt[, c("base", "pred") 
      := data.table(stri_split_fixed(text, " ", simplify = TRUE))]
    dt <- dt[, text := NULL]
    dt <- setkey(dt, base)
    }
    
valid.pentagram <- split_base_pred(ngram = pentagram)
rm(pentagram, corpus.sentences)

# further prepare validation data for entry into prediction algorithm
valid.pentagram.base <- valid.pentagram[["base"]]
valid.pentagram.pred <- valid.pentagram[["pred"]]

```
```{r Evaluate model}
# function to run validation data through model
predict_next_word <- function(prev.words, model = ngrams.freq.15){
    string <- str_replace_all(prev.words, "_", " ")
    n = 4:1
    patt = sprintf("\\w+( \\w+){0,%d}$", n-1)
    test.input <- data.table(base = stri_extract(string, regex = patt))
    test.input <- test.input[, order := length(test.input$base):1]
    prep.input <- test.input[, base := str_replace_all(test.input$base, "\\s", "_") %>% 
        tolower()] 
    prep.input <- setkey(prep.input, base)
    prep.input <- prep.input$base
    results.dt <- model[base %in% prep.input]
    result <- results.dt[order(-prob)[1], .(base, pred)]
    print(result)
    }

ptm <- proc.time() # start time
predict_next_word(prev.words = valid.pentagram.base[58912])
proc.time() - ptm # end time and difference


num.to.test <- 1:1000
model.test <- map_df(.x = valid.pentagram.base[num.to.test], .f = predict_next_word) %>%
    mutate(actual.pred = valid.pentagram.pred[num.to.test],
           result = pred == actual.pred)
model.test %>% summarise(correct.pred.rate = sum(result[!is.na(result)])/length(result),
                         no.pred = sum(is.na(result))/length(result))

ptm <- proc.time() # start time
ngrams.freq.15[base == "given_that"]
proc.time() - ptm # end time and difference

```


# Thoughts and todo:
- learn how to use multiple cores (parallel processing)
- learn how to index / create dictionary to speedup processing
need to remove bad words... perhaps I can group by sentences then remove sentences with bad words
- create unigram, bigram, trigram, tetragram, pentagram
- split into base and predict words
- determine frequencies for each base / predict word pair
- combine all data and save to file
- create program to run predictions using backoff method. start at 5 grams then end with most common unigram work if no solution is found



```{r week 3 quiz}
# 1. The guy in front of me just bought a pound of bacon, a bouquet, and a case of
# soda
# beer - answer good
# cheese
# pretzels

# 2. You're the reason why I smile everyday. Can you follow me please? It would mean the
# universe
# most
# best
# world - answer good

# 3. Hey sunshine, can you follow me and make me the
# saddest
# smelliest
# happiest - answer no
# bluest

# 4. Very early observations on the Bills game: Offense still struggling but the
# crowd
# players
# referees
# defense - answer no

# 5. Go on a romantic date at the
# mall
# beach - answer no
# grocery
# movies

# 6. Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my
# phone
# way - answer close (would work if I dropped count under 2)
# horse
# motorcycle

# 7. Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some
# weeks
# thing
# time - answer yes (better if I drop count of 1)
# years

# 8. After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little
# toes
# eyes
# ears
# fingers - answer no

# 9. Be grateful for the good times and keep the faith during the
# sad
# worse
# bad - answer no
# hard

# 10. If this isn't the cutest thing you've ever seen, then you must be
# callous
# insensitive
# asleep
# insane - answer no


```



```{r tests}
ngram_fcm <- function(tokens, ngrams) {

  ngms <- tokens

  # get rid of tokens metadata not necessary for our UC
  ngms_lst <-  as.list(ngms)
  ngms_unlst  <- unlist(ngms_lst) # (named) character with _ sep. ngrams

  # split in " "-separated pairs:  "n-1 tokens", "nth token"
  ngms_blank_sep <- stringi::stri_replace_last_fixed(ngms_unlst,"_", " ")

  # list of character(2)  ( (n-1)gram ,nth token )
  tk2_lst <- tokens(ngms_blank_sep)

  # --- end of tokens/ngrams pre-processing

  # ordinary fcm
  fcm_ord <- fcm(tk2_lst , ordered = TRUE)
}

bigram.fcm <- ngram_fcm(tokens = corpus.test.bigram, ngrams = 2)
trigram.fcm <- ngram_fcm(corpus = corpus.test.sentences, ngrams = 3)

sample_code()
[1] "based on https://github.com/quanteda/quanteda/issues/1413#issuecomment-414795832"
[1] "great package great support, thanks"
Feature co-occurrence matrix of: 7 by 6 features.
7 x 6 sparse Matrix of class "fcm"
         features
features  a b 1 2 3 4
  3_a_b_2 0 0 0 0 1 0
  a_b_2_3 0 0 0 0 0 1
  b_2_3_4 1 0 0 0 0 0
  2_3_4_a 0 1 0 0 0 0
  3_4_a_b 0 0 0 0 1 0
  4_a_b_3 0 0 0 0 0 1
  a_b_3_4 0 0 0 0 0 0   



bigram.dfm <- dfm(corpus.test.bigram)

text.input <- "for the following reasons I want to"

str_extract(text.input, "[[:alpha:]]* [[:alpha:]]*$")

test.text <- c(sample.blogs, sample.news, sample.twitter)

test.corpus <- corpus(test.text)

tokens.test <- tokens(test.corpus, remove_punct = TRUE)

token.dfm <- dfm(corpus.test.unigram)

unigram.freq <- textstat_frequency(token.dfm)

bi.dfm <- dfm(corpus.test.bigram)
bigram.freq <- textstat_frequency(bi.dfm)

sum.token.dfm <- colSums(token.dfm)

token.dfm.prop <- dfm_weight(token.dfm, scheme = "prop")
sum.token.dfm.prop <- colSums(token.dfm.prop)

system.time(bigram.dfm <- tokens_ngrams(tokens.test, n = 2))

# removes features with less than 10 occurances
dfm_trim(dfm, min_termfreq = 10)
# converts counts to proportions
dfm_weight(dfm, scheme  = "prop")
#  translates dictionary values to keys in a DFM.
dfm_lookup()

# example
news_dfm <- dfm(corp, remove = stopwords('en'), remove_punct = TRUE)
news_dfm <- dfm_remove(news_dfm, pattern = c('*-time', 'updated-*', 'gmt', 'bst'))
news_dfm <- dfm_trim(news_dfm, min_termfreq = 100)


test <- c("b_c", "d_c", "g_h")
dt.test <- data.table(text = test)
dt.split.test <- dt.test[, c("base", "pred")
                         := data.table(stri_split_fixed(test, "_",
                                                        simplify = TRUE))]

data.table(stri_split_fixed(test, "_", simplify = TRUE))

```
# Ideas to improve accuracy and efficiency
- 

```{r dfm tests}
text.test <- c("Hello my friend how are you doing today.", "Please come to my office to recieve your pay check. That would be a good time.", "Give me a call when you have the chance.", "Hello my friend how are you doing today.", "Please come to my office to recieve your pay check. That would be a good time.", "Give me a call when you have the chance.", "Give me a call when you have the chance.")

# Prepare corpus by sentences
corpus.sentences <- corpus(text.test) %>%
    corpus_reshape(., to = "sentence") 

# bigrams
# start by initiating all cores for parallel processing
bigram <- create_ngrams(corpus = corpus.sentences, ngrams = 2L)

dfm.bigram <- dfm(bigram)

dfm.colsums <- colSums(dfm.bigram)

dt.bigram <- data.table(features = featnames(dfm.bigram), counts = dfm.colsums)

split_base_pred_freq <- function(ngram){
    dt <- data.table(text = ngram)
    dt <- dt[, text := stri_replace_last_fixed(text,"_", " ")]
    dt <- dt[, c("base", "pred") 
      := data.table(stri_split_fixed(text, " ", simplify = TRUE))]
    dt <- dt[, text := NULL]
    dt <- setkey(dt, base)
    dt <- dt[, ":=" (count = uniqueN(pred),
               base.count = .N),
             key = base]
    dt <- dt[count > 1]
    dt <- dt[, prob := count / base.count]
    dt <- dt[ , .SD[which.max(count)],
              key = base]
    return(dt)
}
```

