---
title: "Next_word_pred_model"
author: "Joe Rubash"
date: "December 12, 2018"
output: html_document
---


# plan of attack ("make it work", "make it right", "make it fast")
Stage one - make a simple model using calculated frequencies from n-grams
1. read in data and make a corpus with all sources
2. to start out sample a small portion of this corpus
3. use qunteda package to create unigram (to lower case, strip punc, remove extra 
white space. etc)
4. create a dictionary to assign integer to unique words in unigram
5. build unigram, bigram and trigram document term matrices
6. split ngrams to give previous word(s) and next word (to predict)
7. calc frequencies
8. to colsum of frequencies to
9. create backoff method to start with highest order n-gram then move to lowest
order n-gram if no solution found
10. create method to evaluate accuracy and efficiency of model. Make this easy to 
run so I can later itterate quickly through different ideas to improve


stage two use smoothing to improve prediction and various methods to spee up
8. figure out how to alter frequencies to predict unkowns
9. su

```{r setup}
# clear environment
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)

# required libraries
library(tidyverse)
library(readr)
library(readtext)
library(purrr)
library(quanteda)
library(tm)
library(data.table)
library(stringi)
library(stringr)
library(knitr)
library(rbenchmark)

ptm <- proc.time()
# Download data-----------------------------------------------------------------
# download and unzip data folder
if(!dir.exists("../data/Coursera-SwiftKey/")){
    file.url <- "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
    temp <- tempfile()
    download.file(file.url, temp, mode = "wb")
    unzip(temp, exdir = "../data/Coursera-SwiftKey")
    rm(temp)
}

# Sample data------------------------------------------------------------------
# define sample size, read in all english data files, sample, combine, add lines and save to file as csv
sample.size <- 0.15 # delete sample files and rerun code if changing sample size

# create function to read in convert, sample then write to file text data
sample_data <- function(file){
    readLines(con = file,
              skipNul = TRUE,
              encoding="UTF-8") %>%
    sample(.,
           size = sample.size * length(.),
           replace = FALSE) #%>% 
        #data.table(text = .)
    }

# read in and sample all text data
sample.blogs <- sample_data(file = "../data/Coursera-SwiftKey/final/en_US/en_US.blogs.txt")
sample.news <- sample_data(file = file("../data/Coursera-SwiftKey/final/en_US/en_US.news.txt", "rb"))
sample.twitter <- sample_data(file =  "../data/Coursera-SwiftKey/final/en_US/en_US.twitter.txt")

# Combine all text samples
sample.text <- c(sample.blogs, sample.news, sample.twitter) 

# remove unneeded objects
rm(sample.blogs, sample.news, sample.twitter)

# the following code seems to work well to create unigram and bigram tokens from 
# combined text. The tokens can then be converted from lists to a single character
# vector.

# Prepare corpus by sentences---------------------------------------------------
corpus.sentences <- corpus(sample.text) %>%
    corpus_reshape(., to = "sentence") 

# Functions to create ngrams and then calculate frequencies---------------------
# Create ngrams
create_ngrams <- function(corpus, ngrams){
    tokens(corpus, what = "word",
                              remove_numbers = TRUE,
                              remove_punct = TRUE,
                              remove_symbols = TRUE,
                              remove_separators = TRUE,
                              remove_twitter = TRUE, 
                              remove_hyphens = TRUE,
                              remove_url = TRUE,
                              ngrams = ngrams,
                              skip = 0L,
                              concatenator = "_",
                              verbose = quanteda_options("verbose"),
                              include_docvars = TRUE) %>% 
    tokens_tolower(.) %>% 
    unlist(., use.names = FALSE)
    }

# Model data
# convert to data.table then split bigram into base and prediction
split_base_pred_freq <- function(ngram){
    dt <- data.table(text = ngram)
    dt <- dt[, text := stri_replace_last_fixed(text,"_", " ")]
    dt <- dt[, c("base", "pred") 
      := data.table(stri_split_fixed(text, " ", simplify = TRUE))]
    dt <- dt[, text := NULL]
    dt <- setkey(dt, base)
    dt <- dt[, ":=" (count = uniqueN(pred),
               base.count = .N),
             key = base]
    dt <- dt[, prob := count / base.count]
    dt <- dt[ , .SD[which.max(count)],
              key = base]
    return(dt)
}

# Create ngram frequencies data table-------------------------------------------

# unigrams
unigram <- create_ngrams(corpus = corpus.sentences, ngrams = 1L)
unigram.dt <- data.table(text = unigram)
unigram.dt <- unigram.dt[, count := .N, by = text]
unigram.dt <- unigram.dt[, prob := count / length(unigram.dt$count)]
unigram.dt <- unigram.dt[order(count, decreasing = TRUE)]
unigram.dt <- unique(unigram.dt)

# bigrams
# start by initiating all cores for parallel processing
bigram <- create_ngrams(corpus = corpus.sentences, ngrams = 2L)
bigram.freq <- split_base_pred_freq(ngram = bigram)
rm(bigram)

# trigrams
trigram <- create_ngrams(corpus = corpus.sentences, ngrams = 3L)
trigram.freq <- split_base_pred_freq(ngram = trigram)
rm(trigram)

# quadgrams
quadgram <- create_ngrams(corpus = corpus.sentences, ngrams = 4L)
quadgram.freq <- split_base_pred_freq(ngram = quadgram)
rm(quadgram)

# pentagrams
pentagram <- create_ngrams(corpus = corpus.sentences, ngrams = 5L)
pentagram.freq <- split_base_pred_freq(ngram = pentagram)
rm(pentagram, corpus.sentences)

# combine ngrams
ngram.list <- list(bigram.freq, trigram.freq, quadgram.freq, pentagram.freq)
ngrams.freq <- rbindlist(ngram.list)
rm(ngram.list, bigram.freq, trigram.freq, quadgram.freq, pentagram.freq)
ngrams.freq <- setkey(ngrams.freq, base)

proc.time() - ptm

```
```{r Run model}

ptm <- proc.time()
ngrams.freq[base == "give_me"]
proc.time() - ptm

benchmark(ngrams.freq[base == "give_me"])

# start of methods to input base words into model to predict next word
text.input <- "been close"
# extract last two words from input text
trim.input <- str_extract(text.input, "[[:alpha:]]* [[:alpha:]]*$") %>% 
    tolower() %>% 
    str_replace_all(., "\\s", "_")

blogs.trigram.freq[base == trim.input, pred]


#test <- "the beach"
#test <- tolower(test)
#last4 <- word(string, start = -4, end = -1)
#last3 <- word(string, start = -3, end = -1)
#last2 <- word(string, start = -2, end = -1)
#last1 <- word(string, -1)
#
#test.words <- c(last4, last3, last2, last1) %>% 
#    str_replace_all(., "\\s", "_")

ngrams.freq[base == "case_of"]
```


# Thoughts and todo:
- learn how to use multiple cores (parallel processing)
- learn how to index / create dictionary to speedup processing
need to remove bad words... perhaps I can group by sentences then remove sentences with bad words
- create unigram, bigram, trigram, tetragram, pentagram
- split into base and predict words
- determine frequencies for each base / predict word pair
- combine all data and save to file
- create program to run predictions using backoff method. start at 5 grams then end with most common unigram work if no solution is found




```{r tests}
ngram_fcm <- function(tokens, ngrams) {

  ngms <- tokens

  # get rid of tokens metadata not necessary for our UC
  ngms_lst <-  as.list(ngms)
  ngms_unlst  <- unlist(ngms_lst) # (named) character with _ sep. ngrams

  # split in " "-separated pairs:  "n-1 tokens", "nth token"
  ngms_blank_sep <- stringi::stri_replace_last_fixed(ngms_unlst,"_", " ")

  # list of character(2)  ( (n-1)gram ,nth token )
  tk2_lst <- tokens(ngms_blank_sep)

  # --- end of tokens/ngrams pre-processing

  # ordinary fcm
  fcm_ord <- fcm(tk2_lst , ordered = TRUE)
}

bigram.fcm <- ngram_fcm(tokens = corpus.test.bigram, ngrams = 2)
trigram.fcm <- ngram_fcm(corpus = corpus.test.sentences, ngrams = 3)

sample_code()
[1] "based on https://github.com/quanteda/quanteda/issues/1413#issuecomment-414795832"
[1] "great package great support, thanks"
Feature co-occurrence matrix of: 7 by 6 features.
7 x 6 sparse Matrix of class "fcm"
         features
features  a b 1 2 3 4
  3_a_b_2 0 0 0 0 1 0
  a_b_2_3 0 0 0 0 0 1
  b_2_3_4 1 0 0 0 0 0
  2_3_4_a 0 1 0 0 0 0
  3_4_a_b 0 0 0 0 1 0
  4_a_b_3 0 0 0 0 0 1
  a_b_3_4 0 0 0 0 0 0   








bigram.dfm <- dfm(corpus.test.bigram)

text.input <- "for the following reasons I want to"

str_extract(text.input, "[[:alpha:]]* [[:alpha:]]*$")

test.text <- c(sample.blogs, sample.news, sample.twitter)

test.corpus <- corpus(test.text)

tokens.test <- tokens(test.corpus, remove_punct = TRUE)

token.dfm <- dfm(corpus.test.unigram)
unigram.freq <- textstat_frequency(token.dfm)

bi.dfm <- dfm(corpus.test.bigram)
bigram.freq <- textstat_frequency(bi.dfm)

sum.token.dfm <- colSums(token.dfm)

token.dfm.prop <- dfm_weight(token.dfm, scheme = "prop")
sum.token.dfm.prop <- colSums(token.dfm.prop)

system.time(bigram.dfm <- tokens_ngrams(tokens.test, n = 2))

# removes features with less than 10 occurances
dfm_trim(dfm, min_termfreq = 10)
# converts counts to proportions
dfm_weight(dfm, scheme  = "prop")
#  translates dictionary values to keys in a DFM.
dfm_lookup()

# example
news_dfm <- dfm(corp, remove = stopwords('en'), remove_punct = TRUE)
news_dfm <- dfm_remove(news_dfm, pattern = c('*-time', 'updated-*', 'gmt', 'bst'))
news_dfm <- dfm_trim(news_dfm, min_termfreq = 100)


test <- c("b_c", "d_c", "g_h")
dt.test <- data.table(text = test)
dt.split.test <- dt.test[, c("base", "pred")
                         := data.table(stri_split_fixed(test, "_",
                                                        simplify = TRUE))]

data.table(stri_split_fixed(test, "_", simplify = TRUE))

```
# Ideas to improve accuracy and efficiency
- 
```{r}
sample_code <- function() {

  require(quanteda)

  print(paste("based on","https://github.com/quanteda/quanteda/issues/1413#issuecomment-414795832"))
  print("great package great support, thanks")

  ngms <- tokens("a b 1 2 3 a b 2 3 4 a b 3 4 5", n = 2:5)

  # get rid of tokens metadata not necessary for our UC
  ngms_lst <-  as.list(ngms)
  ngms_unlst  <- unlist(ngms_lst) # (named) character with _ sep. ngrams

  # split in " "-separated pairs:  "n-1 tokens", "nth token"
  ngms_blank_sep <- stringi::stri_replace_last_fixed(ngms_unlst,"_", " ")

  # list of character(2)  ( (n-1)gram ,nth token )
  tk2_lst <- tokens(ngms_blank_sep)

  # --- end of tokens/ngrams pre-processing

  # ordinary fcm
  fcm_ord <- fcm(tk2_lst , ordered = TRUE)

  fcm_ord[33:39, 1:6]
}


sample_code()
[1] "based on https://github.com/quanteda/quanteda/issues/1413#issuecomment-414795832"
[1] "great package great support, thanks"
Feature co-occurrence matrix of: 7 by 6 features.
7 x 6 sparse Matrix of class "fcm"
         features
features  a b 1 2 3 4
  3_a_b_2 0 0 0 0 1 0
  a_b_2_3 0 0 0 0 0 1
  b_2_3_4 1 0 0 0 0 0
  2_3_4_a 0 1 0 0 0 0
  3_4_a_b 0 0 0 0 1 0
  4_a_b_3 0 0 0 0 0 1
  a_b_3_4 0 0 0 0 0 0   
```

