---
title: "Next Word Prediction App"
author: "Joe Rubash"
date: "January 12, 2019"
output: 
  ioslides_presentation: 
    smaller: no
    widescreen: no
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# clear enviroment
rm(list = ls())

# load required libraries
library(shiny)
library(data.table)
library(stringi)
library(stringr)
library(dplyr)
library(tidyr)
library(purrr)
library(magrittr)

# load model
ng.model <- readRDS(file = "./shiny.presentation.data/ngrams.mod.4.clean.rds")

# load validation data
valid.hexagram.base <- readRDS(file = "./shiny.presentation.data/valid.hexagram.base.rds")
valid.hexagram.pred <- readRDS(file = "./shiny.presentation.data/valid.hexagram.pred.rds")

# function to run validation data through model
predict_next_word <- function(prev.words, model = ng.model){
    string <- str_replace_all(prev.words, "_", " ")
    n = 4:1
    patt = sprintf("\\w+( \\w+){0,%d}$", n-1)
    test.input <- data.table(base = stri_extract(string, regex = patt))
    test.input <- test.input[, order := length(test.input$base):1]
    prep.input <- test.input[, base := str_replace_all(test.input$base, "\\s+", "_") %>% 
        tolower()] 
    prep.input <- setkey(prep.input, base)
    prep.input <- prep.input$base
    results.dt <- model[base %in% prep.input]
    result <- results.dt[order(-prob)[1], .(base, pred)]
    print(result)
    }
```

## Modivation for the app | Why would we want to predict the next word?  
* We spend a lot of time typing  
* If an app could predict the next word you intent to type it could 
speed up the typing process.   
* Benefits could include:  
    + Making apps easier to use => good user experience  
    + Allow those with disabilities to more easily communicate
    + Correct common typing mistakes mistakes (duplicate words)
    


## Building The Prediction Model  
### General strategy:  
A large collection of text was used to determine the probability that a word would 
occur given the previous word or words. Using this probability table, just the last 
few words of new text could be used to find the associated next word with the highest
probability.

### A few model building details:  
* Text from twitter, blogs and news feeds were used to build the model    
* Removing punctuation and non ASCII characters helped to improve accuracy   
* Final probabilities were adjusted to give greater weight to longer ngrams

```{r validation, message=FALSE, warning=FALSE, include=FALSE}
# validation of models 

# randomly sample from validation data set
valid.size <- 1000
seed <- 562
set.seed(seed)
valid.base.sample <- sample(valid.hexagram.base, valid.size, replace = FALSE)

# associated actual prediction for sample of validation data
set.seed(seed)
valid.pred.sample <- sample(valid.hexagram.pred, valid.size, replace = FALSE)

ptm <- proc.time() # start time
model.test <- map_df(.x = valid.base.sample, .f = predict_next_word) %>%
    mutate(actual.pred = valid.pred.sample,
           result = pred == actual.pred)
succes.rate <- model.test %>% summarise(correct.pred.rate = sum(result[!is.na(result)])/length(result),
                         no.pred = sum(is.na(result))/length(result))
# percent success rate
perc.success.rate <- succes.rate[[1]]*100
# time per test
time.per.test <- round((proc.time() - ptm)[3] / valid.size, 4) # end time and difference

```

## Application of The Model
<style>
  .col1 {
    columns: 1 400px;
    -webkit-columns: 1 400px;
    -moz-columns: 1 400px;
  }
  .col2 {
    columns: 1 400px;         /* number of columns and width in pixels*/
    -webkit-columns: 1 400px; /* chrome, safari */
    -moz-columns: 1 400px;    /* firefox */
  }
</style>

<div class="col2">
To use the prediction model, the last 5, 4, 3, 2 and 1 words are taken from the 
input text. These word combinations are then found in the model and the associated
next word with the highest probability is returned as the predicted next word.  
</div>

### Performance
From `r valid.size` tests (using text never seen by the model)

<div class="col1">
* Correctly predicted the next word `r perc.success.rate`% of the time  
* Took roughly `r time.per.test` seconds per test  
* The model occupies `r round(object.size(ng.model)/1000000, 1)`MB in the R environment 
</div> 
 
----
```{r next word app, message=FALSE, warning=FALSE, ehco=FALSE}

shinyApp(
    ui = fluidPage(
        titlePanel(
        h2("Next Word Prediction App")),
        # instructions
        h3('Type in the box below and the next word will be predicted when you click the SPACEBAR. Click ENTER to combine into a sentence.'),
        h4("This app is intended to make life a little easier by predicting the 
next word you intend to type based on previous words. If the predicted word matches
your expectations hit ENTER and all text will be saved and then displayed below.
Continue to follow this method and a long string of text will be created fairly quickly."),
        
        # script to allow "enter" key to be used instead of pressing button
        tags$script(' $(document).on("keydown", function (e) {
                                     Shiny.onInputChange("lastkeypresscode", e.keyCode);
                                     });
                                     '),
        
        # text input box for previous words
        textInput("previous.words", "Previous words:"),
        #div(style="display: inline-block;vertical-align:top; width: 250px;",textInput("previous.words", "Previous words:")),
        # text output giving predicted next word based on previous
        textInput("inText", "Predicted next word:"),
        #div(style="display: inline-block;vertical-align:top; width: 100px;",textInput("inText", "Next word:")),          
        textAreaInput("combination", "Combination:", width = '300px', height = '60px'),
        actionButton("clear.text", "Start over")
    ),

server = function(input, output, session) {
    # code below observes text in previous word input box then automatically runs
    # through prediction model to give output
    
    #functions for later use:-------------------------------------------------------
    # save predicted text data and combine with past data
    saveData <- function(data) {
        if (exists("textString")) {
            textString <<- paste(textString, data)
        }else{
            textString <<- data
        }
    }
    
    # load updated text data
    loadData <- function() {
        if (exists("textString")) {
            textString <- str_replace_all(textString, "\\s+", " ")
            textString
        }
    }
    
    # code below observes text in previous word input box then runs through
    # prediction model when the space bar is used
    
    # Predict next word --------------------------------------------------------
    # Run previous words through model to give predicted next word
    observe({
        if(!is.null(input$lastkeypresscode)){
            if(input$lastkeypresscode == 32){
                # previous words input
                x <- input$previous.words
                # function to run previous word string through model
                predict_next_word <- function(prev.words, model = ng.model){
                    # remove spaces in front and behind word
                    prev.words <- trimws(prev.words)
                    # remove punctuation
                    prev.words <- str_replace_all(prev.words, "[:punct:]", "")
                    # break into 5 to 1 word groups
                    n = 5:1
                    patt = sprintf("\\w+( \\w+){0,%d}$", n-1)
                    test.input <- data.table(base = stri_extract(prev.words, regex = patt))
                    #test.input <- test.input[, order := length(test.input$base):1]
                    # replace spaces with "_" to give same format as model
                    prep.input <- test.input[, base := str_replace_all(test.input$base, "\\s+", "_") %>%
                                                 tolower()] 
                    #prep.input <- setkey(prep.input, base)
                    # isolate just base words
                    prep.input <- prep.input$base
                    # find predictions in model
                    results.dt <- model[base %in% prep.input]
                    # take result with highest probability
                    result <- results.dt[order(-prob)[1], pred][1]
                    # what to do for no result
                    result <- ifelse(is.null(prev.words), "",
                                     ifelse(is.na(result), "prediction", result))
                    print(result)
                }
                
                prediction <- predict_next_word(prev.words = x)
                # This will change the value of input$inText, based on x
                
                updateTextInput(session,
                                "inText",
                                value = paste(prediction))
            }
        }
    })
    
    # Combine previous and predicted 
    # function to save previous words and prediction
    textData <- reactive({
        data <- paste(input$previous.words, input$inText)
        #data <- data.table(returned.text = data)
    })
    
    # When the "Yes, that's it!" button is clicked, save the data
    observe({
        if(!is.null(input$lastkeypresscode)){
            if(input$lastkeypresscode == 13){
                # save combination of previous and predicted text
                saveData(textData())
                updateTextAreaInput(session,
                                    "combination",
                                    value = loadData())
            }
        }
    })
    
    # Store and display output ---------------------------------------------
    
    # clear text boxes to allow another prediction to be made
    observe({
        if(!is.null(input$lastkeypresscode)){
            if(input$lastkeypresscode == 13){
                # clear content of previous words and predicted text boxes
                updateTextInput(session, "previous.words", value = "")
                updateTextInput(session, "inText", value = "")}
        }
    })
    
    # clear stored text data------------------------------------------------
    # Button to clear stored text data and start over
    observeEvent(input$clear.text, {
        # clear saved data
        textString <<- NULL
        updateTextAreaInput(session,
                            "combination",
                            value = loadData()
        )
    },ignoreNULL = TRUE, ignoreInit = TRUE, once = FALSE)
}
    
)

```