---
title: "Next_word_pred_model"
author: "Joe Rubash"
date: "December 12, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---



# Model description:
I am having a hard time building this model with current RAM. Even after separating ngram creation I am still running into RAM limit at quadgrams. 25% sample of text seems to be the limit for me at the moment.

This model is similar to model 4 but uses a training data set that is 50% of all text data. RAM is reduced by saving each aggrigated ngram to file to combined later. Probability for each base-pred pair is modified slightly based on the length of ngrams. This is based off the "stupid back off model". I will multiply each probability by 0.4 raised to the absolute value of 5 - number of words in the base. This equation is assuming a max base of 5 words which would see no adjustment.

```{r setup}
# clear environment
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE)

# required libraries
library(tidyverse)
library(readr)
library(readtext)
library(purrr)
library(quanteda)
library(tm)
library(data.table)
library(reshape2)
library(stringi)
library(stringr)
library(knitr)
library(rbenchmark)
library(doParallel)
library(R.utils)

quanteda_options(threads = 3)

# Functions---------------------------------------------------------------------
  
# function to create dfm from ngrams of specied length
dfm_ngrams <- function(corpus, ngrams){
    dfm(corpus, what = "word",
        tolower = TRUE,
        remove_numbers = TRUE,
        remove_punct = TRUE,
        remove_symbols = TRUE,
        remove_separators = TRUE,
        remove_twitter = TRUE, 
        remove_hyphens = TRUE,
        remove_url = TRUE,
        ngrams = ngrams,
        skip = 0L,
        concatenator = "_",
        verbose = quanteda_options("verbose")) %>% 
        colSums(.)
    }

# Model data
# convert to data.table then split bigram into base and prediction and calc probabilities
split_base_pred_freq <- function(dfm.ngram){
    #dfm.ngram <- bigrams.test
    dt <- data.table(feats = names(dfm.ngram), count = as.data.table(dfm.ngram))
    dt <- dt[, feats := stri_replace_last_fixed(feats,"_", " ")]
    dt <- dt[, c("base", "pred") 
      := data.table(stri_split_fixed(feats, " ", simplify = TRUE))]
    dt <- dt[, feats := NULL]
    dt <- setkey(dt, base)
    dt <- dt[, `:=` (base.count = sum(count.dfm.ngram),
                     ngram = length(str_split(base, "_")[[1]])),
             key = base]
    dt <- dt[count.dfm.ngram > 1]
    dt <- dt[, `:=` (prob = (count.dfm.ngram / base.count) * (0.4^abs(5-ngram)))]
    dt <- dt[ , .SD[which.max(count.dfm.ngram)],
              key = base]
    dt <- dt[, count.dfm.ngram := NULL]
    dt <- dt[, base.count := NULL]
    dt <- dt[str_which(base, "[[:ascii:]]+")] # extract only ascii characters
    dt <- dt[-(str_which(base, "[:digit:]"))] # remove words with numbers
    return(dt)
}

```
```{r load and sample data}
# download and unzip data folder if needed
if(!dir.exists("../data/Coursera-SwiftKey/")){
    file.url <- "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
    temp <- tempfile()
    download.file(file.url, temp, mode = "wb")
    unzip(temp, exdir = "../data/Coursera-SwiftKey")
    rm(temp)
}

ptm <- proc.time() # start time
# Sample and slpit into train, validate and test data

# read in all text data
blogs <- readLines(con = "../data/Coursera-SwiftKey/final/en_US/en_US.blogs.txt",
                   skipNul = TRUE,
                   encoding="UTF-8")
news <- readLines(con = file("../data/Coursera-SwiftKey/final/en_US/en_US.news.txt", "rb"),
                  skipNul = TRUE,
                  encoding="UTF-8")
twitter <-  readLines(con = "../data/Coursera-SwiftKey/final/en_US/en_US.twitter.txt",
                      skipNul = TRUE,
                      encoding="UTF-8")

# Combine all text data
all.text <- c(blogs, news, twitter) 

# remove unneeded objects
rm(blogs, news, twitter)

sample.size <- 0.50 # delete sample files and rerun code if changing sample size
valid.size <- 0.05
test.size <- 0.05

# split all.text into training, validation and test
set.seed(123)
valid <- sample(all.text,
                   size = valid.size * length(all.text),
                   replace = FALSE)
test <- sample(all.text[-which(valid %in% all.text)],
                   size = test.size * length(all.text),
                   replace = FALSE) 
training <- sample(all.text[-which(c(valid, test) %in% all.text)],
                   size = sample.size * length(all.text),
                   replace = FALSE) 

# remove unneeded objects
rm(all.text)
```
```{r corpus and bigram}
# Prepare corpus by sentences---------------------------------------------------
corpus.sentences <- corpus(training) %>%
    corpus_reshape(., to = "sentence") 
# save to file to be used later
saveRDS(corpus.sentences, file = "./data/temp.data/corpus.sentences.rds")
rm(training)

# Create ngram frequencies data table-------------------------------------------

# unigrams
#unigram <- create_ngrams(corpus = corpus.sentences, ngrams = 1L)
#unigram.dt <- data.table(text = unigram)
#unigram.dt <- unigram.dt[, count := .N, by = text]
#unigram.dt <- unigram.dt[, prob := count / length(unigram.dt$count)]
#unigram.dt <- unigram.dt[order(count, decreasing = TRUE)]
#unigram.dt <- unique(unigram.dt)

# bigrams
# start by initiating all cores for parallel processing
bigram <- dfm_ngrams(corpus = corpus.sentences, ngrams = 2) %>%
    split_base_pred_freq(.)
# save to file to be combined later
saveRDS(bigram, file = "./data/temp.data/bigram.rds")
rm(bigram)
```
```{r trigram}
# restart the R session to clear RAM
.rs.restartR() 
# run library and functions script "setup"

#load corpus.sentences
corpus.sentences <- readRDS(file = "./data/temp.data/corpus.sentences.rds")

# trigrams
trigram <- dfm_ngrams(corpus = corpus.sentences, ngrams = 3) %>%
    split_base_pred_freq(.)
# save to file to be combined later
saveRDS(trigram, file = "./data/temp.data/trigram.rds")
rm(trigram)
```
```{r quadgram}
# restart the R session to clear RAM
.rs.restartR() 
# run library and functions script "setup"

#load corpus.sentences
corpus.sentences <- readRDS(file = "./data/temp.data/corpus.sentences.rds")

# quadgrams
quadgram <- dfm_ngrams(corpus = corpus.sentences, ngrams = 4) %>%
    split_base_pred_freq(.)
# save to file to be combined later
saveRDS(quadgram, file = "./data/temp.data/quadgram.rds")
rm(quadgram)
```
```{r pentagram}
# restart the R session to clear RAM
.rs.restartR() 
# run library and functions script "setup"

#load corpus.sentences
corpus.sentences <- readRDS(file = "./data/temp.data/corpus.sentences.rds")

# pentagrams
pentagram <- dfm_ngrams(corpus = corpus.sentences, ngrams = 5) %>%
    split_base_pred_freq(.)
# save to file to be combined later
saveRDS(pentagram, file = "./data/temp.data/pentagram.rds")
rm(pentagram)
```
```{r hexagram and combine}
# restart the R session to clear RAM
.rs.restartR() 
# run library and functions script "setup"

#load corpus.sentences
corpus.sentences <- readRDS(file = "./data/temp.data/corpus.sentences.rds")

# hexagrams
hexagram <- dfm_ngrams(corpus = corpus.sentences, ngrams = 6) %>%
    split_base_pred_freq(.)
rm(corpus.sentences)

# load ngrams
bigram <- readRDS(file = "./data/temp.data/bigram.rds")
trigram <- readRDS(file = "./data/temp.data/trigram.rds")
quadgram <- readRDS(file = "./data/temp.data/quadgram.rds")
pentagram <- readRDS(file = "./data/temp.data/pentagram.rds")

# combine ngrams
ngram.list <- list(bigram, trigram, quadgram, pentagram, hexagram)
ngrams.freq <- rbindlist(ngram.list)
rm(ngram.list, bigram, trigram, quadgram, pentagram, hexagram)
ngrams.freq <- setkey(ngrams.freq, base)

# Save an object to file under the "next_word_prediction" app
saveRDS(ngrams.freq, file = "./Next_word_prediction/model.data/ngrams.mod.5.rds")

proc.time() - ptm # end time and difference
# Restore the object
#ng.15.test <- readRDS(file = "./Next_word_prediction/model.data/ngrams.freq.15.rds")

```
Time to run above model 1-9-2019 11 pm
  user  system elapsed 
2290.42  195.03 2512.83 


# Thoughts and todo:
- learn how to use multiple cores (parallel processing)
- learn how to index / create dictionary to speedup processing
need to remove bad words... perhaps I can group by sentences then remove sentences with bad words
- create unigram, bigram, trigram, tetragram, pentagram
- split into base and predict words
- determine frequencies for each base / predict word pair
- combine all data and save to file
- create program to run predictions using backoff method. start at 5 grams then end with most common unigram work if no solution is found



